%% -*- Mode:TeX; Fonts:(hl12fb) -*-
%% *-* File: /usr/local/gbbopen/doc-source/acknowledgments.tex *-*
%% *-* Last-Edit: Mon Mar 21 04:13:39 2011; Edited-By: cork *-*
%% *-* Machine: twister.local *-*

%% Copyright (C) 2003-2008, Dan Corkill <corkill@GBBopen.org>
%% Part of the GBBopen Project.
%% Licensed under Apache License 2.0 (see LICENSE for license information).

%% ========================================================================
%%  Shared Acknowledgments text

\section*{Acknowledgments}
\T\addcontentsline{toc}{section}{\textbf{Acknowledgments}}%
\label{sec:acknowledgments}%
\index{acknowledgments}%

Many people have contributed comments, suggestions, design ideas, questions
(and answers), bug reports, and code to GBBopen, and we appreciate their time
and effort.  Acknowledgment of some of their contributions here does not
necessarily imply that any individual or the organizations with which they are
affiliated endorse GBBopen or this documentation.  Disclaimers aside, GBBopen
users thank each of you!

Douglas Crosher ported GBBopen to \xsitelink{Scieneer
  CL}{http://www.scieneer.com/scl/}.  \xsitelink{Gary
  King}{http://www.metabang.com/} worked on the initial \xsitelink{Digitool
  MCL}{http://www.digitool.com/} and
\xsitelink{OpenMCL}{http://openmcl.clozure.com/} porting efforts.
Christian Lynbech performed the initial CMUCL port.  \xsitelink{Sam
  Steingold}{http://www.podval.org/~sds/} initiated the
\xsitelink{CLISP}{http://clisp.cons.org/} port.
Vladimir Tzankov provided Portable Threads support for \xsitelink{CLISP}{http://clisp.cons.org/}/MT.

Questioners, bug reporters, capability demanders, contributors, and great idea
suggesters include: \xsitelink{Pascal Costanza}{http://p-cos.net/}, Matthew
Danish, Michael Hannemann, Susan Lander, Attila Lendvai, Wendall Marvel,
\xsitelink{Clayton Morrison}{http://www.cs.arizona.edu/~clayton/}, Beryl
Nelson, Eric O'Connor, \xsitelink{Zack
  Rubinstein}{http://www.cs.cmu.edu/~zbr/}, Bill St. Clair, Earl Wagner, Paul
Werkowski, and \xsitelink{Huzaifa Zafar}{http://www.cs.umass.edu/~hzafar/}.

\paragraph{Organizations}

\xsitelink{Franz Inc.}{http://www.franz.com} and \xsitelink{LispWorks
  Ltd.}{http://www.lispworks.com} provided (and continue to provide)
Common Lisp licenses and technical support to the Project.

Some early design work for GBBopen was supported by DARPA's Information
Exploitation Office (\xsitelink{IXO}{http://dtsn.darpa.mil/ixo/}) under
contract \nobr{MDA-972-02-C-0028} to \xsitelink{Information Extraction \&
  Transport, Inc.}{http://www.iet.com/}

Other efforts using GBBopen that have indirectly led to contributed GBBopen
improvements and enhancements include: research supported by the ``Fusion
Based Knowledge for the Future Force'' ATO program and the ``Advanced REsearch
Solutions - Fused Intelligence with Speed and Trust'' program at the U.S. Army
RDECOM CERDEC Intelligence and Information Warfare Directorate, Fort Monmouth,
NJ, under contract \nobr{W15P7T-05-C-P621}; research on ``Command \& Control
and Data Fusion Architectures'' supported by DND Canada under contract
\nobr{W7701-4-2118}; work on ``Knowledgable Dynamic-Process Modeling and
Execution'' supported by Boeing and Infosys Technologies Limited; research on
``A Multi-Agent Approach for Heterogeneous Persistent Surveillance'' supported
by Raytheon Intelligence \& Information Systems; work on ``Massive-Scale
Representation and Reasoning'' supported by GHX; and research supported by the
AFRL ``Advanced Computing Architecture'' program, under contract
\nobr{FA8750-05-1-0039}.

\paragraph{Legacy Contributors}

GBBopen builds upon concepts and ideas that were explored and refined in the
UMass Generic Blackboard system and the commercial GBB product.  The following
people made significant contributions to those systems:

\begin{tabular}{llll}
  ~~~~~& \textbf{UMass Generic} & ~~~~~~~~~~~~~ & \textbf{GBB Product}\\
  & \textbf{Blackboard System} & & Tony Carrico \\  
  & Dan Corkill & & Dan Corkill \\
  & Kevin Gallagher & & Raymond de Lacaze\\
  & Philip Johnson & & Kevin Gallagher\\
  & Kelly Murray & & Susan Lander \\
  & & & Zack Rubinstein\\
  & & & Suzanne Tromara\\
\end{tabular}

The UMass Generic Blackboard Project received research support from The
National Science Foundation, the Defense Advanced Research Projects Agency,
the Office of Naval Research, and Texas Instruments, Inc.

%% ========================================================================
